Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-12 WSSSPE1_report/WSSSPE1_summary_report.tex
Refine abstract.

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
This document summarizes the First Workshop on Sustainable Software for Science:
Practice and Experiences (WSSSPE1). The content is based on the workshop
itself, as recorded in a set of collaborative notes taken during the workshop,
including discussion about the keynote presentations, three themed panels that
discussed papers accepted at the workshop, other discussion, cross-cutting
issues, use cases, and conclusions. It joins a nascent literature seeking to
understand what drives software work in science and how the reward systems of
science work, thereby shaping the type of software work undertaken, including the
extent to which developers are motivated to build software for the long-term,
for the use of others, and whether to work collaboratively or separately.
Unique perspectives were captured about issues such as comprehensive documentation,
software deployment, and development practices; software licenses; career paths
for developers; identifiers for software through a
mechanism such as a Digital Object Identifier or publication of a ``software
paper''; and evidence of software contribution and impact, recorded in online
systems such as source code repositories like GitHub. For truly sustainable
software, there should be no endpoint, as the software products continue to be
used and useful beyond a single institution, grant, and developer or
development team.

@@

Because progress in scientific research is dependent on the quality of and
accessibility to software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments who are an integral and embedded part of
successful research teams. The WSSSPE workshop was intended to provide a forum
for discussion of the challenges, including both positions and experiences. The
short papers and debates have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this
document. An estimate of 90 to 150 participants were present
throughout the day. Additional papers based on extended versions of workshop
submissions are expected. The level of interest in the workshop has led the
organizers, working with some of the submitters and attendees, to plan two
additional workshops (at the 2014 SciPy and SC14 conferences), to obtain
funds from the National Science Foundation and the Gordon and Betty Moore
Foundation to support these workshops, and to turn the
workshop website into a community website that can be used as a focus for
further discussion and progress. Additionally, a minisymposium that aims to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 Society for Industrial and Applied Mathematics
(SIAM) Annual Meeting on ``Reliable
Computational Science.''

>>>>

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
This document summarizes the First Workshop on Sustainable Software for Science:
Practice and Experiences (WSSSPE1). The content is based on 
a set of collaborative notes taken during the workshop,
including discussion about the keynote presentations, three themed panels that
discussed papers accepted at the workshop, other discussion, cross-cutting
issues, use cases, and conclusions. It joins a nascent literature seeking to
understand what drives software work in science and how the reward systems of
science function, thereby shaping the types of software projects undertaken, including the
extent to which developers are motivated to build software for the long-term,
for the use of others, and whether to work collaboratively or separately.
Unique perspectives were captured about issues such as comprehensive documentation, 
development and deployment practices; software licenses; career paths
for developers; identifiers for software through a
mechanism such as a Digital Object Identifier or publication of a ``software
paper''; and evidence of software contribution and impact, recorded in online
systems such as source code repositories like GitHub. For truly sustainable
software, there should be no endpoint, as the software products continue to be
used and useful beyond a single institution, grant, and developer or
development team.

@@

Because progress in scientific research is dependent on the quality of and
accessibility to software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments as an integral part of
successful research teams. The WSSSPE workshop was intended to provide a forum
for discussion of the challenges, including both positions and experiences. The
short papers and debates have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this
document. An estimate of 90 to 150 participants were present
throughout the day. Additional papers based on extended versions of workshop
submissions are expected. The level of interest in the workshop has led the
organizers, working with some of the submitters and attendees, to plan two
additional workshops (at the 2014 SciPy and SC14 conferences), to obtain
funds from the National Science Foundation and the Gordon and Betty Moore
Foundation to support these workshops, and to turn the
workshop website into a community website that can be used as a focus for
further discussion and progress. Additionally, a minisymposium that aims to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 Society for Industrial and Applied Mathematics
(SIAM) Annual Meeting on ``Reliable
Computational Science.''
